The SMT Competition 2015--2018

### 2018-10-09 2018/report/SMT-COMP.tex
Goals, FLoC ranking

The SMT Competition is organized under the direction of the SMT
Steering Committee, which appoints the competition chair.  This
appointment should happen ``within 30 days from the end of the current
edition of the SMT workshop''~\cite{smt-bylaws} but in practice has
often happened later.  It is then up to the competition chair to
assemble a team of organizers, and to distribute the work for the next
competition.

Each year, the SMT Workshop includes a block of time to present the
competitors and results of the competition.  The workshop is
affiliated with a different conference each year: with CAV in 2015,
with IJCAR in 2016, again with CAV in 2017, and with FLoC in~2018.
Consequently, its date varies by several weeks.  Competition results
were presented on July~19 in 2015, on July~2 in 2016, on July~23 in
2017, and on July~13 in 2018.

@@

Around the initial solver deadline, the organizers also aim to publish
the latest version of the competition tools, and--in collaboration
with the SMT-LIB maintainers--to release a new version of the SMT
Library.

  Also: Call for Comments in February/March\\
  Draft rules in April

  SMT-LIB release dates:\\
  2015-06-01 (vs. 06-01)\\
  2016-05-23 (vs. 05-15)\\
  2017-06-05 (vs. 06-04)\\
  2018-05-20 (vs. 05-27)
\end{tjark}

@@

By number of solver versions submitted, the competitions in 2015--2018
were the four largest in the history of SMT-COMP.  On average, 26
solver versions were submitted each year since 2015.  In contrast, the
competitions from 2005--2014 only received an average of 12
submissions~\cite{smtcomp-2014}.

@@

In 2014, the SMT Competition was part of the FLoC Olympic Games.
These gave the competition increased visibility and sponsored three
Kurt G\"odel medals.  After substantial internal debate, the 2014
competition organizers decided to award these to the winner of the
main track's QF\_BV division (which has historically been the logic
with the largest number of solver submissions), as well as to the two
best main-track solvers according to a competition-wide ranking that
had been instituted especially for this
purpose~\cite{DBLP:journals/jsat/CokDW14}.  This ranking emphasizes
breadth of solver participation---a solver participating in many
logics need not be the best in any one of them to rank highly.

The competition-wide ranking for the main track was retained and
refined after~2014.  A contentious issue in this ranking has been the
scoring of incorrect solver responses.  In the per-division rankings,
an incorrect response causes a solver to be ranked behind all correct
solvers~\cite{rules18}.  But the competition's main track consists of
many independent divisions.  To balance the strong interest in correct
solvers against the risk of stifling innovation, the competition-wide
ranking was revised in~2017 to use a fixed numeric penalty for each
division with incorrect responses.  The value of this penalty is
chosen so that entering a (possibly buggy) solver that can solve all
benchmarks into a division has a positive expected score if the
probability of a soundness bug in the solver is below~$20\,\%$.

The 2018 edition of the competition was again part of the FLoC Olympic
Games.  After last-minute negotiations with the Olympic Games
organizers, it was decided that the SMT Competition would award seven
Kurt G\"odel medals.  The 2018 competition organizers conferred these
to the following solvers:
\begin{itemize}
\item The best three main-track solvers according to the
  competition-wide scoring (first place: CVC4, second place:
  Yices-2.6.0, third place: SMTInterpol);
\item the solver(s) that won the QF\_BV division in the main track
  (sequential ranking: Boolector, parallel ranking: Minkeyrink-MT);
\item the solver that won the most competitive divisions in the
  application track (Yices-2.6.0);
\item the solver that won the most competitive divisions in the
  unsat-core track (Yices-2.6.0).
\end{itemize}

\begin{tjark}
  Discuss competition-wide rankings for application and unsat-core
  tracks.
\end{tjark}

@@

>>>>

The SMT Competition was instituted in 2005.  It has been held annually
since then (with the exception of 2013, when an
evaluation~\cite{DBLP:journals/jar/CokSW15} was conducted instead of a
competition).  The main goals of the competition were to spur adoption
of the common, community-designed SMT-LIB formats (see
Section~\ref{sec:smtlib}), and to spark further advances in SMT,
especially for verification~\cite{smtcomp2005}.  These, together with
providing a useful yardstick for users and developers of SMT solvers,
are still its main goals today.  The competition has been successful
in establishing SMT-LIB as the \emph{de facto} standard language for
SMT solvers.  In recent years, the focus in this regard has shifted
towards promoting some of the newer extensions of the SMT-LIB
language, such as a floating-point theory~\cite{TODO} or algebraic
datatypes~\cite{TODO}.

The competition is organized under the direction of the SMT Steering
Committee, which appoints the competition chair.  This appointment
should happen ``within 30 days from the end of the current edition of
the SMT workshop''~\cite{smt-bylaws} but in practice has often
happened later.  The competition chair then assembles a team of
organizers, and distributes the work for the next competition.
%
The SMT Workshop includes a block of time to present the competitors
and results of the competition.  The workshop is affiliated with a
different conference each year: with CAV in 2015, with IJCAR in 2016,
again with CAV in 2017, and with FLoC in~2018.  Consequently, its date
varies by several weeks.  Competition results were presented on
July~19 in 2015, on July~2 in 2016, on July~23 in 2017, and on July~13
in~2018.

@@

Around the initial solver deadline, the organizers also aim to publish
the latest version of the competition tools, and---in collaboration
with the SMT-LIB maintainers---to release a new version of the SMT
Library.  SMT-LIB releases were made on June~1 in 2015, on May~23 in
2016, on June~5 in 2017, and on May~20 in 2018.  Anyone may submit new
benchmarks to the SMT Library, and competitors are permitted to tune
their solvers accordingly.  Therefore it is important that all
benchmarks eligible for the competition are released at least some
time before the final solver submission deadline.  To give the SMT-LIB
maintainers sufficient time for curation, the deadline for new
benchmark contributions has usually been about four to six weeks
before the initial solver deadline, typically in April or early May.

The competition rules are revised each year, and a (near final) draft
version of the rules is made public on the competition web site
typically around mid-April.  The competition chair has ultimate
responsibility for the rules, but changes are often preceded by
discussions among the organizers or within the SMT community.  Such
discussions are initiated by a Call for Comments that is typically
sent out to the SMT-COMP mailing list~\cite{TODO} in February or early
March.  Publicly, this call is the first harbinger of a new edition of
the SMT Competition.

@@

By number of solver versions submitted, the competitions in 2015--2018
were the four largest in the history of SMT-COMP.  On average, 26
solver versions were submitted each year since 2015.  In contrast, the
competitions from 2005--2014 only received an average of 12
submissions~\cite{DBLP:journals/jsat/CokDW14}.

@@

In 2014, the SMT Competition was part of the FLoC Olympic Games.
These gave the competition increased visibility and sponsored three
Kurt G\"odel medals.  After substantial internal debate, the 2014
competition organizers decided to award these to the winner of the
main track's QF\_BV division (which has historically been the logic
with the largest number of solver submissions and job pairs), as well
as to the two best main-track solvers according to a competition-wide
ranking that had been instituted especially for this
purpose~\cite{DBLP:journals/jsat/CokDW14}.  This ranking emphasizes
breadth of solver participation---a solver participating in many
logics need not be the best in any one of them to rank highly.

The competition-wide ranking for the main track was retained and
refined after~2014.  A contentious issue with this ranking has been
the scoring of incorrect solver responses.  In the per-division
rankings, an incorrect response causes a solver to be ranked behind
all correct solvers~\cite{rules18}.  But the competition's main track
consists of many independent divisions.  To balance the strong
interest in correct solvers against the risk of stifling innovation,
the competition-wide ranking was revised in~2017 to use a fixed
numeric penalty for each division with incorrect responses.  The value
of this penalty is set so that entering a (possibly buggy) solver that
can solve all benchmarks into a division has a positive expected score
if the probability of a soundness bug in the solver is below~$20\,\%$.

The 2018 edition of the competition was again part of the FLoC Olympic
Games.  After last-minute negotiations with the Olympic Games
organizers, it was decided that the SMT Competition would award seven
Kurt G\"odel medals.  The 2018 competition organizers conferred these
to the following solvers:
\begin{itemize}
\item The best three main-track solvers according to the
  competition-wide ranking (first place: CVC4, second place:
  Yices-2.6.0, third place: SMTInterpol);
\item the solver(s) that won the QF\_BV division in the main track
  (sequential ranking: Boolector, parallel ranking: Minkeyrink-MT);
\item the solver that won the most competitive divisions in the
  application track (Yices-2.6.0);
\item the solver that won the most competitive divisions in the
  unsat-core track (Yices-2.6.0).
\end{itemize}

In comparison to the competition-wide ranking for the main track, the
metric used for the application and unsat-core track---simply counting
the number of competitive divisions that a solver has won---was fairly
unsophisticated.  Future editions of the competition should perhaps
refine this to use a ranking scheme that is more similar to the
main-track ranking, by weighing divisions according to their relative
importance, and by also rewarding good performance by solvers other
than division winners.  A difficulty in adapting the main-track
ranking for the other tracks is that each main-track benchmark is
worth one point (before weighing) if solved correctly, while
benchmarks in the application and unsat-core track are worth a varying
number of points.  Moreover, in the unsat-core track, the size of the
smallest unsatisfiable core---and hence the number of points
attainable on any given benchmark---may not be known precisely.

@@

\TODO
